---
title: Linearity for Tranduction Efficiency Assay Qualification
authors:
  - name: Adeline Chew
    affiliation: Tikva Allocell Pte. Ltd.
    roles: writing
    corresponding: false
  - name: Yu Xin Lim
    affiliation: Tikva Allocell Pte. Ltd.
    roles: writing
    corresponding: false
  - name: Jinkai Teo
    affiliation: Tikva Allocell Pte. Ltd.
    roles: writing
    corresponding: true
    email: jinkaiteo@tikvaallocell.com
    orcid: 0009-0005-0663-2118
bibliography: 
  - references.bib
  - grateful-refs.bib
number-sections: true
---

## Introduction

This report documents the statistical analysis performed for the linearity of the Transduction Efficiency (TE) Assay. This document is intended to be read as part of the qualification report for Tranduction Efficiency. In brief, the TE assay employs a flow cytometry method to establish the proportion of transduced cells in the test article (cell suspension). In the manufacturing of the CART cells, B7-H3 and Serpin B9 (CAS) were introduced into the T cells via transduction using a replication-incompetent retrovirus vector. Fluorescent-labeled antibodies targeting the VHH and T2A were used to label the B7-H3.CAR and the transduced Serpin B9 (CAS). The proportion of the positively stained cells (B7-H3 and Serpin B9 (CAS) positive) out of the total amount of cells in the cell suspension is reported as the Transduction Efficiency.

```{r packages, echo=FALSE, warning=FALSE, output = FALSE}
library(tidyverse)
library(kableExtra)
library(grateful)
library(jtools)
library(huxtable)
library(broom)
library(flextable)
library(webshot2)
cite_packages(out.dir = ".")
```

## Data generated for Linearity

The IHC harmonized tripartite guideline [@Borman_Elder_2017] defines linearity as:

::: {.callout-note title="IHC Guidlines"}
A linear relationship should be evaluated across the range of the analytical procedure. It may be demonstrated directly on the drug substance (by dilution of a standard stock solution) and/or separate weighings of synthetic mixtures of the drug product components, using the proposed procedure. The latter aspect can be studied during investigation of the range.
:::

In line with this guidance, samples with different transduction efficient were generated using CART cells expressing the B7-H3 CAR antigen (drug substance, DS) are diluted at various known ration (volume per volume) with untransduced T cells that does not express the CAR on the surface, as shown in @tbl-rawdata. The expected transduction efficiency was normalized based on the 100%(v/v) and 0 %(v/v). These set of samples were ran by two operators across three days to incorporate inter-operator and inter-day variations.

```{r data}
rawdata <- "Operator	Day	day-mean	Expected
Operator 1	Day1	83.62	84.55055556
Operator 1	Day2	84.16	84.55055556
Operator 1	Day3	84.73	84.55055556
Operator 2	Day1	84.26333333	84.55055556
Operator 2	Day2	83.25	84.55055556
Operator 2	Day3	87.28	84.55055556
Operator 1	Day1	69.7	67.64111111
Operator 1	Day2	67.39333333	67.64111111
Operator 1	Day3	68.44	67.64111111
Operator 2	Day1	68.34	67.64111111
Operator 2	Day2	68.26333333	67.64111111
Operator 2	Day3	68.48	67.64111111
Operator 1	Day1	52.28	50.73166667
Operator 1	Day2	51.54	50.73166667
Operator 1	Day3	49.99	50.73166667
Operator 2	Day1	52.43	50.73166667
Operator 2	Day2	50.61	50.73166667
Operator 2	Day3	49.11	50.73166667
Operator 1	Day1	33.49333333	33.82222222
Operator 1	Day2	33.33	33.82222222
Operator 1	Day3	33.43	33.82222222
Operator 2	Day1	33.79333333	33.82222222
Operator 2	Day2	32.43	33.82222222
Operator 2	Day3	32.38	33.82222222
Operator 1	Day1	16.85	16.91277778
Operator 1	Day2	14.81333333	16.91277778
Operator 1	Day3	13.65	16.91277778
Operator 2	Day1	16.84	16.91277778
Operator 2	Day2	14.94	16.91277778
Operator 2	Day3	13.7	16.91277778
Operator 1	Day1	0	0.003333333
Operator 1	Day2	0	0.003333333
Operator 1	Day3	0.013333333	0.003333333
Operator 2	Day1	0	0.003333333
Operator 2	Day2	0	0.003333333
Operator 2	Day3	0.006666667	0.003333333
"
tab_data <- read.table(text = rawdata, sep = "\t", header = TRUE)

tab_data <- tab_data %>% dplyr::arrange(Operator, Day, Expected)
```

```{r rawdata}
#| label: tbl-rawdata
#| tbl-cap: "Data generated for linearity in qualification"

tab_data %>% kableExtra::kbl(col.names = c(
  "Operator", "Day", "Observed (%TE)", "Expected (%TE)"
), digits = 2, align = "c") %>% 
  kableExtra::collapse_rows(column = c(1, 2)) %>%
  kable_classic_2()
```

## Methods

Taking reference to the IHC harmonized tripartite guideline [@Borman_Elder_2017]:

::: {.callout-note title="IHC Guidlines"}
Linearity should be evaluated by visual inspection of a plot of signals as a function of analyte concentration or content. If there is a linear relationship, test results should be evaluated by appropriate statistical methods, for example, by calculation of a regression line by the method of least squares. In some cases, to obtain linearity between assays and sample concentrations, the test data may need to be subjected to a mathematical transformation prior to the regression analysis. Data from the regression line itself may be helpful to provide mathematical estimates of the degree of linearity.

The correlation coefficient, y-intercept, slope of the regression line and residual sum of squares should be submitted. A plot of the data should be included. In addition, an analysis of the deviation of the actual data points from the regression line may also be helpful for evaluating linearity.
:::

A similar method was employed in this analysis where a linear model is used for the regression of the data collected within each operator within the same day.

```{r citingPKG}
cite_packages(output = "paragraph", out.dir = ".")
```

## Results
```{r ssError}
regression_list <- list()
anova_list <- data.frame()

for (i in unique(tab_data$Operator)) {
  
  i_data <- tab_data %>% dplyr::filter(Operator == i)
  
  for (j in unique(i_data$Day)) {
    
    j_data <- tab_data %>% dplyr::filter(Operator == i, Day == j)
    reg_ij <- stats::lm(data = j_data, formula = day.mean ~ Expected)
    SSmodel_ij <- stats::anova(reg_ij) %>% cbind(Operator = i, Day = j)
    SSmodel_ij$Terms <- rownames(SSmodel_ij)
    rownames(SSmodel_ij) <- NULL
    anova_list <- anova_list %>% rbind(SSmodel_ij)
    regression_list[[paste(i, j, collapse = ".", sep = ", ")]] <- reg_ij
    
  }
}

anova_list_model <- anova_list %>% dplyr::filter(Terms == "Expected") %>% dplyr::select(Operator, Day, `Sum Sq`, `F value`, `Pr(>F)`) %>% dplyr::rename(Model = `Sum Sq`)
anova_list_Error <- anova_list %>% dplyr::filter(Terms == "Residuals") %>% dplyr::select(Operator, Day, `Sum Sq`) %>% dplyr::rename(Error = `Sum Sq`)

anova_list_SS <- anova_list_Error %>% dplyr::left_join(anova_list_model, by = c("Operator", "Day")) %>% dplyr::mutate(Total = Error + Model) %>% 
  select(Operator, Day, Error, Model, Total, `F value`, `Pr(>F)`)
```

The y-intercept, slope and R2 are summarized in @tbl-reg, while the sum of squares are summarized in @tbl-ss.

```{r regression}
#| label: tbl-reg
#| tbl-cap: "Regression parameters for linearity results"

export_summs(regression_list)
```

```{r sum-of-squares}
#| label: tbl-ss
#| tbl-cap: "Sum of squares for linearity results"

anova_list_SS %>% kableExtra::kbl(digits = 2, align = "c") %>% 
  kableExtra::collapse_rows(column = c(1)) %>%
  add_header_above(c(" ", " ", "Sum of Squares" = 3, "Model Fit" = 2)) %>% 
  kable_classic_2()
```

@fig-regression summarizes in a plot, the various linearity samples across the two operators and 3 days.

```{r linearRegression}
#| label: fig-regression
#| fig-cap: "Linear regression of linearity samples within each operator across the three days."

ggplot2::ggplot(data = tab_data, aes(x = Expected, y = day.mean, group = Operator)) + 
  ggplot2::geom_smooth(method = "lm", aes(color = "red")) +
  ggplot2::geom_point(aes(shape = Day)) +
  ggplot2::facet_grid(cols = vars(Operator)) +
  ggplot2::labs(x = "Expected TE%", y = "Observed TE%") + 
  ggplot2::theme_light() + 
  guides(fill = "none", color = "none", linetype = "none")
```

## Conclusion

The data from this set of data support the claim of linearity of the Transduction Efficiency (TE) Assay over the range from 0 to 80% TE.


## References {.unnumbered}

:::{#refs}

:::